Work Fragmentation in Developer Interaction Data

### 2016-01-27 paper_2.tex
modifications in RQ5

A more detailed description of each category follows.
\RR{check category description, this is done from my (flaky) memory}

@@

Category \emph{file} regroup events related to file management, such as opening and saving files. \RR{more details here, not sure} This kind of activity is indicative of program editing in the sense of saving files \RR{not quite sure here}

@@

Category \emph{debug} is concerned with events that occurr during debugging sessions, such as setting and removing breakpoints, stepping in the source code, going backwards/forwards, stepping into the execution of a method, etc. This category is clearly related to program understanding.

Category \emph{text-nav} regroups events related to text-level navigation, such as moving the cursor left, right, to the beginning/end of lines, or up/down a line. If this category includes navigation activity, it is still related to low-level artifacts as it represents navigation over single files, and usually small distances (characters or lines). On the contrary, category \emph{high-nav}, is related to the navigation between more abstract program elements (classes, methods, packages, interfaces), rather than low-level source code. As such, \emph{text-nav} is more closely related to code editing, while \emph{high-nav} is more closely related to program understanding.

Category \emph{search} and \emph{find-replace} are both concerned with search activities, although there are subtle differences. \emph{search} regroups most kinds of search tools, such as text search, or searching code entities (types, packages, methods). As such, it is more to be seen as a high-level program navigation tool (the search results may span several files and are presented in a view that allows quick access to the entities---using such a view is classified in the \emph{high-nav} category). On the other hand, \emph{find-replace} concerns commands that are usually employed for text navigation and more particularly to perform editions (systematically replacing occurences of a string with another). If the first category is related to program exploration, the second is more related to program edition. There is a caveat here, as it is possible to use \emph{find-replace} commands only to find elements, which is more indicative of program exploration; however, the UDC dataset does not allow us to distinguish these usages.

Category \emph{perspective} deals with events related to perspective change in Eclipse, such as switching from the coding to the debugging perspective, or vice-versa \RR{other perspectives too}. This is more indicative of ... \RR{what?}

@@

Category \emph{refactoring} regroups usage of refactorings tools (rename, move methods, extract methods, etc). This is indicative of edition activities \RR{expand}

Category \emph{clean-build} regroup events related to build system usage (build the system, clean the build, run the build, the tests, etc). This is indicative of \RR{out of inspiration and time for now ...}

@@

\paragraph{Comparison with observational studies}
Our theoretical foundations are based on observational studies performed on information workers' development environments \cite{MGH05, IH07, LVD06, PD10}. These findings helped us understand the response to work fragmentation according to the recorded user activity, and it was possible to made a relation between the user's data and the phenomena that occur on the working places describe on the studies.

@@

We first analyzed the Recovery Time in detail, finding a higher productivity when the interruption is positive and we identified a positive evolution of the metrics as the programmer advances through this stage, reaching a peak of productivity between the last minutes of the Recovery Time and 10 minutes before the next interruption happens.

In addition, we found that after a positive interruption the most common events during the first minutes of the Recovery Time are those with an assertive intention such as edition, refactoring and saving files. In contrast, after a negative interruption the most common events are debugging, searching and navigation around classes, which tell us about a doubtful and uncertain behavior from the programmer. These findings support out hypotheses that positive interruptions are due to information-seeking, while negative interruptions are due to context rebuilding after an actual interruption.

>>>>

A more detailed description of each category follows.

@@

Category \emph{file} regroup events related to file management, such as opening and saving files, refreshing its contents and editing its properties. This kind of activity is indicative of program editing in the sense of (mostly) saving files.

@@

Category \emph{debug} is concerned with events that occur during debugging sessions, such as setting and removing breakpoints, stepping in the source code, going backwards/forwards, stepping into the execution of a method, etc. This category is clearly related to program understanding.

Category \emph{text-nav} regroups events related to text-level navigation, such as moving to the next word, to the beginning/end of lines, or up/down a line. If this category includes navigation activity, it is still related to low-level artifacts as it represents navigation over single files, and usually small distances (characters or lines). On the contrary, category \emph{high-nav}, is related to the navigation between more abstract program elements (classes, methods, packages, interfaces), rather than low-level source code. As such, \emph{text-nav} is more closely related to code editing, while \emph{high-nav} is more closely related to program understanding.

Category \emph{search} and \emph{find-replace} are both concerned with search activities, although there are subtle differences. \emph{search} regroups most kinds of search tools, such as text search, or searching code entities (types, packages, methods). As such, it is more to be seen as a high-level program navigation tool (the search results may span several files and are presented in a view that allows quick access to the entities---using such a view is classified in the \emph{high-nav} category). On the other hand, \emph{find-replace} concerns commands that are usually employed for text navigation and more particularly to perform editions (systematically replacing occurrences of a string with another). If the first category is related to program exploration, the second is more related to program edition. There is a caveat here, as it is possible to use \emph{find-replace} commands only to find elements, which is more indicative of program exploration; however, the UDC dataset does not allow us to distinguish these usages.

Category \emph{perspective} deals with events related to perspective change in Eclipse, such as switching from the coding to the debugging perspective, or vice-versa. Eclipse allows the user to program in different environments, like enterprise, mobile and Web. It also supports activities regarding other kind of roles, like testing, design and database management. Each of this environments and roles might require to open a different perspective, so having events regarding a perspective change is an indicative of the different roles that a user has.

@@

Category \emph{refactoring} regroups usage of refactoring tools (rename, move methods, extract methods, etc). This is indicative of edition activities in the current piece of code, but it can also affect outer elements and require to verify the changes, implying low and high level navigation.

Category \emph{clean-build} regroup events related to build system usage (build the system, clean the build, run the build, the tests, etc). This kind of events are often automatically executed as part of a grater process. For example, when beginning a debugging session, the IDE might execute a clean and build action at first.

@@

\paragraph{Comparison with observational studies}
Our theoretical foundations are based on observational studies performed on information workers' development environments \cite{MGH05, IH07, LVD06, PD10}. These findings helped us understand the response to work fragmentation according to the recorded user activity, and it was possible to made a relation between the user's data and the phenomena that occur on the working places described in the studies.

@@

We first analyzed the recovery time in detail, finding a higher productivity when the interruption is positive and we identified a positive evolution of the metrics as the programmer advances through this stage, reaching a peak of productivity between the last minutes of the recovery time and 10 minutes before the next interruption happens.

In addition, we found that after a positive interruption the most common events during the first minutes of the recovery time are those with an assertive intention such as edition, refactoring and saving files. In contrast, after a negative interruption the most common events are debugging, searching and navigation around classes, which tell us about a doubtful and uncertain behavior from the programmer. These findings support out hypotheses that positive interruptions are due to information-seeking, while negative interruptions are due to context rebuilding after an actual interruption.
